Convective heat transport in stratified atmospheres at low and high Mach number

### 2016-11-01 tex/nusselt_paper.tex
SUBMISSION DRAFT NUMERO UNO

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convection is essential to heat transport in the cores of high mass stars,
the envelopes of low mass stars, and the atmospheres of terrestrial and 
jovian planets.  The convective dynamics in these astrophysical objects
are influenced by the atmospheric 
stratification, which is small in some systems (e.g. massive star cores)
but spans about
14 density scale heights in the Sun's convective envelope.
Understanding the fundamental
properties of compressible convection in stratified media is essential
to characterizing systems in astrophysics and planetary sciences.  
Numerical constraints have often restricted studies of compressible
convection to moderately high Mach number (Ma), appropriate to regions near 
the Sun's surface.  Few fundamental properties of the 
low-Ma stratified convection, which occurs in the deep solar interior, are known.

Early numerical experiments on stratified convection
in two \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990} and three 
\cite{cattaneo&all1991, brummell&all1996} dimensions
revealed a number of basic properties in the moderate-to-high 
Ma regime. In the widely-studied \RB (hereafter RB) problem, 
upflows and downflows are symmetrical and
the temperature gradient approaches zero in the convective interior 
causing the conductive flux to similarly 
disappear.  These two hallmark characteristics of RB convection change
significantly when stratification is included.  Stratified convection 
exhibits narrow downflow lanes and broad upflow regions.
Furthermore, the \emph{entropy} gradient is negated by convection 
rather than the temperature gradient, such
that in the presence of perfectly efficient convection a significant 
component of the flux is still carried by conduction.

In RB convection, there exist two primary control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity).  These numbers coupled with the aspect ratio of 
the physical domain and the boundary conditions
determine the dynamics of the convection.  In stratified atmospheres, 
in addition to specifying the equation of state and
fundamental properties of the gas, the two control parameters of 
RB convection are joined by the degree of
stratification across the domain and the characteristic 
Ma of the convective flows.  
Polytropically stratified atmospheres, such as those used in 
early studies \cite{graham1975, chan&all1982, hurlburt&all1984, 
cattaneo&all1990, cattaneo&all1991, brummell&all1996} are an ideal extension of
RB convection into the stratified realm as the two additional 
control parameters are directly linked to
basic properties of the atmosphere.  The density stratification is 
set by the number of density scale heights (\nrho)
the atmosphere spans, and Ma is controlled 
by the superadiabatic excess ($\epsilon$),
the deviation of the polytropic index from the adiabatic polytropic 
index \cite{graham1975}.

@@

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We exmine the simplist stratified extension of RB by studying a
fluid composed of monatomic ideal gas particles with an adiabatic 
index of $\gamma = 5/3$ and whose equation of state is $P = R\rho T$. 
This is consistent with the approach used in earlier work.
The initial atmosphere is a plane-parallel polytrope in which 
the gravitational acceleration and conductive flux, 
$\bm{F}_{\text{cond,0}} = -\kappa \grad T_0$, do not vary with depth. To
achieve the latter condition, both $\kappa$ and $\grad T_0$ are constant.
Under these assumptions, satisfying hydrostatic 
equilibrium produces a stratification of
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(z_0 - z)^m, \\
T_0(z)    &= T_{t}(z_0 - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m = m_{\text{ad}} - \epsilon$ is the polytropic index.
The adiabatic polytropic index is $m_{\text{ad}} \equiv (\gamma-1)^{-1}$ and
the superadiabatic excess is $\epsilon$ which sets the scale 
of the entropy gradient ($\grad S_0 \propto -\epsilon$).
A significant advance of this work is the ability to study large 
and small values of $\epsilon$, as will be discussed.
Thermodynamic variables are nondimensionalized at the 
top of the atmosphere as  $P_0(L_z) = \rho_0(L_z) = T_0(L_z) = 1$, 
requiring $z_0 \equiv L_z + 1$ and $R = T_{t} = \rho_{t} = 1$.
By this choice, the non-dimensional length scale is the inverse 
temperature gradient scale and the timescale is the isothermal 
sound crossing time of this unit length.
The height $z$ increases upwards within $[0, L_{z}]$, 
where $L_{z} = e^{n_{\rho}/m} - 1$ is
determined by $n_\rho$ and $\epsilon$.
The characteristic timescale of convective dynamics
is related to the atmospheric buoyancy time, 
$t_{\text{b}} = \sqrt{L_z/g\epsilon}$, with $g = (m+1)$.
Throughout this letter, we use buoyancy time units and 
choose $n_{\rho} = 3$ such that the 
initial density varies by a factor of 20.
All atmospheres studied here have an aspect ratio of 4, 
such that $L_x = 4L_z$.

Atmospheric diffusivities are specified by the Rayleigh number
and the Prandtl number.  The
non-dimensional Rayleigh number is
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0 = \epsilon\ln z_0$ is the entropy difference 
between the top and bottom of the atmosphere, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure, $\nu$ is the kinematic viscosity 
(the viscous diffusivity), and $\chi$ is the thermal diffusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, $\text{Pr} = \nu/\chi$.   
The dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, relate to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  We take
$\mu$ and $\kappa$ to be constant with height.
As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ 
experience an increase in the Rayleigh number 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere. 
In this letter we impose $\text{Pr} = 1$ and specify Ra at 
the top of the domain ($z = L_z$).

At the constant values of $n_\rho$ and Pr used, 
the primary control parameters of convection are $\epsilon$
and Ra.  We decompose our atmosphere into the background 
polytrope ($\ln\rho_{0}, T_{0}$) and the fluctuations
about that background ($\bm{u}, \ln\rho_{1}, T_{1}$),
which can be large.
The scaling of the entropy gradient with $\epsilon$
is reflected in the evolved values of these fluctuations, 
which follow the scaling of
$T_1/T_0 \propto \rho_{1}/\rho_{0} \propto$ Ma$^{2} \propto \epsilon$ 
for low values of $\epsilon$,
as in Fig. \ref{fig:ma_v_eps}.

@@

We evolve the Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\bm{u}$ and adding it to 
(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where the convective flux is
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
and the condcutve flux is $\bm{F}_{\text{cond}} = -\kappa \grad T$.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$, and all four
must all be considered. 
Understanding how these flux interact  
is crucial in characterizing convective heat transport.

@@

The efficiency of convection is quantified by the Nusselt number.  
Nu is well-defined in RB convection
as the total flux normalized by the steady-stateconductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, but one of the earliest definitions used
was \cite{graham1975,hurlburt&all1984}
\begin{equation}
\text{Nu} \equiv \frac{F_{\text{conv, z}} + F_{\text{cond, z}} - F_{\text{A}}}{F_{\text{ref}} - F_{\text{A}}},
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively.  $F_{\text{A}} \equiv -\kappa \partial_z T_{\text{ad}}$ 
is the adiabatic conductive flux and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ 
for an ideal gas in hydrostatic equilibrium.
$F_{\text{ref}} \equiv \Delta T / L_z$, 
where $\Delta T \equiv T(L_z) - T(0)$, is the 
conductive flux of a linear profile connecting the upper
and lower plates, which is constant for 
the choice of fixed-temperature boundaries.

We contend that this is the general form of the Nusselt number
and illustrate this with a few limiting
cases. Convection eliminates
entropy stratification.  Under the Boussinesq approximation,
in which density variations are ignored, entropy stratification 
is directly proportional to temperature stratification
such that $\grad S \rightarrow 0$ when $\grad T \rightarrow 0$.  
The familiar Nu in the RB problem is
therefore retrieved from (\ref{eqn:nusselt}) because $\grad T_{\text{ad}} = 0$.  
In the case of polytropic convection,  $g \rightarrow 0$
as $\epsilon \rightarrow m_{\text{ad}} + 1$.
In this limit, $\grad T_{\text{ad}} \rightarrow 0$ as well and the
definition of the RB Nu is appropriate to use, as convection carries all of the
flux \cite{brandenburg&all2005}. 
However, as $\epsilon \rightarrow 0$, 
$\grad T_{\text{ad}}\rightarrow \grad T_0$ and increasingly
smaller velocity and thermodynamic perturbations are needed to 
achieve $\grad S = 0$ (Fig. \ref{fig:ma_v_eps}a).
These perturbations carry what little flux exists in 
excess of the adiabatic, so the removal of 
$F_{\text{A}}$ in the numerator and denominator of 
(\ref{eqn:nusselt}) is essential for a reasonable Nu measurement.

@@

We evolve initial value problems in which $T_1$ is
filled with infinitesimal, random white noise compared to $T_0$ and $\epsilon$.
We filter the noise spectrum in coefficient space, such that 25\% of the coefficients
have power.
Solutions were time-evolved until a long average of Nu showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection
occurs at $\text{Ra}_c = \{10.06, 10.97, 10.97\}$ for $\epsilon = \{0.5, 10^{-4}, 10^{-7}\}$ respectively.  
We studied Rayleigh
numbers from values at onset up to nearly $10^6$Ra$_c$ for $\epsilon = 0.5$ and for $\epsilon = 10^{-4}$
and up to $10^3$Ra$_c$ for $\epsilon = 10^{-7}$.

At large Ma ($\epsilon = 0.5$), shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}a) once Ra is sufficiently large.
These shocks propagate through upflow upflow regions.  
Such systems were reported in
both two \cite{cattaneo&all1990} and three \cite{malagoli&all1990} dimensional polytropic simulations previously.
These shocks heat material entering the downflows, affecting the dynamics and heat transport
of these systems.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/fluxes_fig.png}
\caption{Time-averaged vertical flux profiles ($F_{\text{z}}$)
for (a) high and (b) low Ma flows at $\text{Ra }= 10^6$.  
All fluxes are defined as in (\ref{eqn:energy_eqn_full}) and 
normalized by $F_{\text{ref}} - F_{\text{A}}$, as in (\ref{eqn:nusselt}).
The dashed lines correspond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  
The grey dash-dot line is the
viscous flux and the green dotted line is the conductive flux with the 
adiabatic contribution removed. Viscous and
potential energy fluxes are negligible in (a) and are not shown.    
The solid black line is
Nu, the properly normalized sum of all the fluxes \label{fig:flux_profiles} }
\end{figure}

High Ma flows at large values of Ra, in which shocks form, exhibit two 
local maxima in the enthalpy flux and kinetic energy flux (Fig. \ref{fig:flux_profiles}a).
Shock-heated fluid parcels sometimes gain vorticity
they sink into the lower atmosphere.
This creates deep, rapidly-rotating regions
of mixing which persist for many overturn times. These ``spinners'' appear to
influence the dynamics, but their contributions are unclear.
At low Ma, only the deep maximum is present (Fig. \ref{fig:flux_profiles}b).  
Fixed-temperature boundary conditions allow the flux at the 
boundaries to vary, so many runs at $\text{Ra }> 10^5$
and $\epsilon = 10^{-4}$ exhibit states in which the flux 
entering the system at the bottom of the atmosphere 
exceeds that which leaves at the top.  
These systems are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  
In these shearing states, the bottom temperature
gradient approaches adiabatic, 
allowing the excess energy to exit through the upper boundary.  
During shearing states, convective transport is suppressed and Nu diminishes.
A proper long-term average over shearing
and non-shearing states retrieves an invariant flux (and Nu) profile throughout the depth
of the atmosphere. 
These shearing states will be covered in more detail in a future paper.

After appropriately time-averaging the fluxes for $\geq 200 t_b$, 
a sensible flux average is retrieved.  Nusselt numbers for
all simulations at low and high Ma are plotted as a function of Ra in Fig. \ref{fig:nu_v_ra}.  
At $\epsilon = 0.5$, in the near-sonic
regime ($\text{Ra} \leq 10^4$), the scaling of Nu with Ra is inflated,
with $\text{Nu} \propto \text{Ra}^{0.45}$, similar to that expected in the
ultimate regime of RB \cite{ahlers&all2009}.  As simulations
pass into the supersonic regime and shocks start to form near the downflows,
that scaling drops to $\text{Nu} \propto \text{Ra}^{0.19}$.  
At $\epsilon = \{10^{-4}, 10^{-7}\}$,
scaling laws of $\text{Nu} \propto \text{Ra}^{\{0.31, 0.31\}}$ are retrieved. 
This scaling is indistringuishable from the $\text{Ra}^{2/7}$ found in RB
convection.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/nu_v_ra.png}
\caption{Variation of Nu as Ra increases at high and low $\epsilon$. 
At high $\epsilon$ (purple circles), 
a clear transition from the subsonic to supersonic regime is evident in the scaling
of Nu with Ra At $\text{Ra} \approx 10^4$.  In the low $\epsilon$ regime (green diamonds and yellow squares), 
our observed Nu scalings collapse onto a similar line which is
indistinguishable from an $\text{Ra}^{2/7}$ scaling, 
as observed in RB convection \cite{johnston&doering2009}.  Error bars are determined
 by the square root of the variance of the time-averaged Nu profile with depth and 
 indicate whether or not a solution is well-converged.
\label{fig:nu_v_ra} }
\end{figure}

>>>>

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convection is essential to heat transport in the cores of high mass stars,
the envelopes of low mass stars, and the atmospheres of terrestrial and 
jovian planets.  The convective dynamics in these astrophysical objects
are influenced by atmospheric 
stratification, which is small in some systems (e.g. massive star cores)
but spans about
14 density scale heights in the Sun's convective envelope.
Understanding the fundamental
properties of compressible convection in stratified media is essential
to characterizing systems in astrophysics and planetary sciences.  
Numerical constraints have often restricted studies of compressible
convection to moderately high Mach number (Ma), appropriate to regions near 
the Sun's surface.  Few fundamental properties of
low Ma convection, which occurs in the deep solar interior, are known.

Early numerical experiments on stratified convection
in two \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990} and three 
\cite{cattaneo&all1991, brummell&all1996} dimensions
revealed a number of basic properties in the moderate-to-high 
Ma regime. In the widely-studied \RB (hereafter RB) problem, 
upflows and downflows are symmetrical and
the temperature gradient approaches zero in the convective interior 
causing the conductive flux to similarly 
disappear.  These two hallmark characteristics of RB convection change
significantly when stratification is included.  Stratified convection 
exhibits narrow downflow lanes and broad upflow regions.
Furthermore, the \emph{entropy} gradient is negated by convection 
rather than the temperature gradient, and
a significant conductive flux can exist in the presence of
efficient convection.

In RB convection, there exist two primary control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity).  These numbers coupled with the aspect ratio of 
the physical domain and the boundary conditions
determine the dynamics of the convection.  In stratified atmospheres, 
in addition to specifying the equation of state and
fundamental properties of the gas, the two control parameters of 
RB convection are joined by the degree of
stratification across the domain and the characteristic 
Ma of the convective flows.  
Polytropically stratified atmospheres, such as those used in 
early studies \cite{graham1975, chan&all1982, hurlburt&all1984, 
cattaneo&all1990, cattaneo&all1991, brummell&all1996}, are an ideal extension of
RB convection into the stratified realm as the two additional 
control parameters are directly linked to
basic properties of the atmosphere.  The density stratification is 
set by the number of density scale heights (\nrho)
the atmosphere spans, and Ma is controlled 
by the superadiabatic excess ($\epsilon$),
the deviation of the polytropic index from the adiabatic polytropic 
index \cite{graham1975}.

@@

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We exmine the simplest stratified extension of RB by studying a
fluid composed of monatomic ideal gas particles with an adiabatic 
index of $\gamma = 5/3$ and whose equation of state is $P = R\rho T$. 
This is consistent with the approach used in earlier work.
The initial atmosphere is a plane-parallel polytrope in which 
the gravitational acceleration and conductive flux, 
$\bm{F}_{\text{cond,0}} = -\kappa \grad T_0$, do not vary with depth. To
achieve the latter condition, both $\kappa$ and $\grad T_0$ are constant.
Under these assumptions, satisfying hydrostatic 
equilibrium produces a stratification of
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(z_0 - z)^m, \\
T_0(z)    &= T_{t}(z_0 - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m = m_{\text{ad}} - \epsilon$ is the polytropic index.
The adiabatic polytropic index is $m_{\text{ad}} \equiv (\gamma-1)^{-1}$ and
the superadiabatic excess is $\epsilon$ which sets the scale 
of the entropy gradient ($\grad S_0 \propto -\epsilon$).
A significant advance of this work is the ability to study large 
and small values of $\epsilon$, as will be discussed.
Thermodynamic variables are nondimensionalized at the 
top of the atmosphere as  $P_0(L_z) = \rho_0(L_z) = T_0(L_z) = 1$, 
requiring $z_0 \equiv L_z + 1$ and $R = T_{t} = \rho_{t} = 1$.
By this choice, the non-dimensional length scale is the inverse 
temperature gradient scale and the timescale is the isothermal 
sound crossing time of this unit length.
The height $z$ increases upwards within $[0, L_{z}]$, 
where $L_{z} = e^{n_{\rho}/m} - 1$ is
determined by $n_\rho$ and $\epsilon$.
The characteristic timescale of convective dynamics
is related to the atmospheric buoyancy time, 
$t_{\text{b}} = \sqrt{L_z/g\epsilon}$, with $g = (m+1)$.
Throughout this letter, we use buoyancy time units and 
choose $n_{\rho} = 3$ such that the 
initial density varies by a factor of 20.
All atmospheres studied here have an aspect ratio of 4, 
such that $L_x = 4L_z$.

The atmospheric diffusivities are primarily controlled by
the non-dimensional Rayleigh number,
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0 = \epsilon\ln z_0$ is the entropy difference 
between the top and bottom boundaries, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure, $\nu$ is the kinematic viscosity 
(the viscous diffusivity), and $\chi$ is the thermal diffusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, $\text{Pr} = \nu/\chi$.   
The dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, relate to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  We take
$\mu$ and $\kappa$ to be constant with height.
As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ 
experience an increase in Ra 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere. 
In this letter we impose $\text{Pr} = 1$ and specify Ra at 
the top of the domain ($z = L_z$).

While holding $n_\rho$ and Pr constant, 
the primary control parameters of convection are $\epsilon$
and Ra.  We decompose our atmosphere into the background 
polytrope ($\ln\rho_{0}, T_{0}$) and the fluctuations
about that background ($\bm{u}, \ln\rho_{1}, T_{1}$),
which can be large.
The scaling of the entropy gradient with $\epsilon$
is reflected in the evolved values of these fluctuations, 
which follow the scaling of
$T_1/T_0 \propto \rho_{1}/\rho_{0} \propto$ Ma$^{2} \propto \epsilon$ 
for low values of $\epsilon$,
as in Fig. \ref{fig:ma_v_eps}.

@@

We evolve the Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\bm{u}$ and adding it to 
(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$, and all four
must all be considered. 
Understanding how these fluxes interact  
is crucial in characterizing convective heat transport.

@@

The efficiency of convection is quantified by the Nusselt number.  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, but one of the earliest definitions used
was \cite{graham1975,hurlburt&all1984}
\begin{equation}
\text{Nu} \equiv \frac{F_{\text{conv, z}} + F_{\text{cond, z}} - F_{\text{A}}}{F_{\text{ref}} - F_{\text{A}}},
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively.  $F_{\text{A}} \equiv -\kappa \partial_z T_{\text{ad}}$ 
is the adiabatic conductive flux and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ 
for an ideal gas in hydrostatic equilibrium.
$F_{\text{ref}} \equiv \Delta T / L_z$, 
where $\Delta T \equiv T(L_z) - T(0)$, is the 
conductive flux of a linear profile connecting the upper
and lower plates, which is constant for 
the choice of fixed-temperature boundaries.

We contend that this is the general form of the Nusselt number
and illustrate this with a few limiting
cases. Convection eliminates
entropy stratification.  Under the Boussinesq approximation,
in which density variations are ignored, entropy stratification 
is directly proportional to temperature stratification
and $\partial_z S \rightarrow 0$ when $\partial_z T \rightarrow 0$.  
The familiar Nu in the RB problem is
therefore retrieved from (\ref{eqn:nusselt}) because $\partial_z T_{\text{ad}} = 0$.  
In the case of polytropic convection,  $g \rightarrow 0$
as $\epsilon \rightarrow m_{\text{ad}} + 1$.
In this limit, $\partial_z T_{\text{ad}} \rightarrow 0$ and the
definition of the RB Nu is appropriate to use \cite{brandenburg&all2005}. 
However, as $\epsilon \rightarrow 0$, 
$\partial_z T_{0}\rightarrow \partial_z T_{\text{ad}}$ and increasingly
smaller velocity and thermodynamic perturbations 
achieve $\partial_z S = 0$ (Fig. \ref{fig:ma_v_eps}a).
These perturbations carry what little flux exists in 
excess of the adiabatic, so the removal of 
$F_{\text{A}}$ in the numerator and denominator of 
(\ref{eqn:nusselt}) is essential for a reasonable Nu measurement.

@@

We evolve initial value problems in which $T_1$ is
filled with infinitesimal, random white noise compared to $T_0$ and $\epsilon$.
We filter the noise spectrum in coefficient space, such that 25\% of the coefficients
have power.
Solutions were time-evolved until a long average of Nu showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection
occurs at $\text{Ra}_c = \{10.06, 10.97, 10.97\}$ for $\epsilon = \{0.5, 10^{-4}, 10^{-7}\}$ respectively.  
We studied Rayleigh
numbers from values at onset up to nearly $10^6$Ra$_c$ for $\epsilon = \{0.5, 10^{-4}\}$ 
and up to $10^3$Ra$_c$ for $\epsilon = 10^{-7}$.

At large Ma ($\epsilon = 0.5$), shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}a) once Ra is sufficiently large.
These shocks propagate through upflow regions.  
Such systems were reported in
both two \cite{cattaneo&all1990} and three \cite{malagoli&all1990} dimensional polytropic simulations previously.
These shocks heat material entering the downflows, affecting the dynamics and heat transport
of these systems.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/fluxes_fig.png}
\caption{Time-averaged vertical flux profiles ($F_{\text{z}}$)
for (a) high and (b) low Ma flows at $\text{Ra }= 10^6$.  
All fluxes are defined as in (\ref{eqn:energy_eqn_full}) and 
normalized by $F_{\text{ref}} - F_{\text{A}}$, as in (\ref{eqn:nusselt}).
The dashed lines correspond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  
The grey dash-dot line is the
viscous flux and the green dotted line is the conductive flux with the 
adiabatic contribution removed. The potential energy flux is negligible and is not shown.
In (a), the viscous flux is negligible and is not shown.  The solid black line is
Nu, the properly normalized sum of all the fluxes \label{fig:flux_profiles} }
\end{figure}

High Ma flows exhibit two 
local maxima in the enthalpy flux and kinetic energy flux (Fig. \ref{fig:flux_profiles}a)
at large enough values of Ra for shocks to form.
Shock-heated fluid parcels sometimes gain vorticity as
they sink into the lower atmosphere.
This creates deep, rapidly-rotating regions
of mixing which persist for many overturn times. These ``spinners'' appear to
influence the dynamics, but their contributions are unclear.

At low Ma, only the deep maximum in enthalpy and kinetic energy fluxes
is present (Fig. \ref{fig:flux_profiles}b).  
Our choice of fixed-temperature boundary conditions allows the flux at the 
boundaries to vary, so many runs at $\text{Ra }> 10^5$
and $\epsilon = 10^{-4}$ exhibit states in which the flux 
entering the system at the bottom of the atmosphere 
exceeds that which leaves at the top.  
These systems are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  
%In these shearing states, flux entering the lower atmosphere approaches $F_\text{A}$,
%allowing excess energy to escape through the upper boundary.  
During shearing states, convective transport is suppressed and Nu diminishes while
excess energy exits the system through the upper boundary.
A proper long-term average over shearing
and non-shearing states retrieves an invariant Nu profile throughout the depth
of the atmosphere. 
These shearing states will be covered in more detail in a future paper.

After appropriately time-averaging the fluxes for $\geq 200 t_b$, 
a sensible flux average is retrieved.  Nusselt numbers for
all simulations at low and high Ma are plotted as a function of Ra in Fig. \ref{fig:nu_v_ra}.  
At $\epsilon = 0.5$, in the near-sonic
regime ($\text{Ra} \leq 10^4$), the scaling of Nu with Ra is inflated,
with $\text{Nu} \propto \text{Ra}^{0.45}$, similar to that expected in the
ultimate regime of RB convection \cite{ahlers&all2009}.  As simulations
pass into the supersonic regime and shocks start to form near the downflows,
that scaling drops to $\text{Nu} \propto \text{Ra}^{0.19}$.  
At $\epsilon = \{10^{-4}, 10^{-7}\}$,
scaling laws of $\text{Nu} \propto \text{Ra}^{\{0.31, 0.31\}}$ are retrieved. 
This scaling is indistinguishable from the $\text{Ra}^{2/7}$ found in RB
convection.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/nu_v_ra.png}
\caption{Variation of Nu as Ra increases at high and low $\epsilon$. 
At high $\epsilon$ (purple circles), 
a clear transition from the subsonic to supersonic regime is evident in the scaling
of Nu with Ra At $\text{Ra} \approx 10^4$.  In the low $\epsilon$ regime (green diamonds and yellow squares), 
our observed Nu scalings collapse onto a similar line which is
indistinguishable from a $\text{Ra}^{2/7}$ scaling, 
as observed in RB convection \cite{johnston&doering2009}.  Error bars are determined
 by the square root of the variance of the time-averaged Nu profile with depth and 
 indicate whether or not a solution is well-converged.
\label{fig:nu_v_ra} }
\end{figure}
